---
title: "Filtering to find IPCC aligned ensemble"
author: "Joe Brown"
date: "2025-03-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goals

This code document filters the Hector + Matilda perturbed parameter ensemble to identify ensemble members  who's central (and upper and lower bounds) are consistent with IPCC AG1-assessed near-, mid-, and long-term warming ranges.

# Set-up

Source files in `set-up` folder.
```{r}
source("set-up/source_all.R")

```

Load model result. 

Model results produced in `Rmd 03` have been saved in the following directory: `output/model_results`.
```{r}
model_result <- readRDS("output/model_results/initial_result_low_ecs_10k.RDS")

```

Confirm that all scenarios retain the same set of ensemble members following NA removal in previous workflow step. This is a check to maintain consistency in downstream analysis.

```{r}
sapply(model_result, nrow)

```

The model result for each scenario should have the same number of rows.

# Process GSAT Data

The following steps extract and normalize the GSAT (`global_tas`) response variable from the model output. We normalize the results to 1995-2014 reference period to prepare the data fro comparison against IPCC WGI assessed warming ranges.  

## Extract GSAT Data

Extract only the `global_tas` variable (Global Surface Air Temperature) from the model results. This is the key variable for for evaluating the ensemble warming range. 

```{r}
# Extract global_tas from each scenario 
gsat_data <- lapply(model_result, function(df) {  
  
  subset_data = subset(df, variable == "global_tas")
  
  return(subset_data)
})

```

## Normalize GSAT to the IPCC reference period (1995-2014)

Normalize the GSAT time series to the 1995-2014 reference period, consistent with IPCC AR6 convention. 

*Note*: The `normalize_to_reference()` function is provided in the project's `utils` folder: `set-up/utils.R`.

```{r}
# Normalize GSAT data to the 1995–2014 reference period
gsat_data_normalized <- lapply(gsat_data, function(df) {
  
  normalized <- normalize_to_reference(df, 
                                       reference_start = 1995, 
                                       reference_end = 2014)
  
  return(normalized)
}) 

```

# Define warming metric and compute warming results

## Define warming metric

Define the metric used to compute *median GSAT for 2081-2100*. This metric will be used to identify ensemble members consistent with IPCC-assessed end-of-century warming ranges. 

*Note*: Additional metrics for other time periods will be computed down stream as needed.

```{r}
# Define end-of-century warming metric (median GSAT over 2081-2100)
warming_metric <- new_metric(var = "global_tas", years = 2081:2100, op = median)
  
```

## Compute warming results 

Compute end-of-century warming values for each ensemble member. These values will be used to identify ensemble members consistent with IPCC-assessed long-term warming ranges.

```{r}
# compute end-of-century warmign results
warming_results <- lapply(gsat_data_normalized, function(df) {
  
  metric_result <- metric_calc(df, warming_metric)
  
  return(metric_result)
})

```

## Check ensemble size

Check the number of available ensemble members for each SSP scenario. This step verifies that all scenarios have the same number of runs, consistent with the precautions implemented in the previous workflow document (`03_run_model_with_low_ecs.Rmd`).

```{r}
# Check the number of runs for each SSP
run_counts <- sapply(warming_results, function(df) nrow(df))

# Print counts to inspect
print(run_counts)

```
All SSPs should have the same number of results.

# Indentify ensemble aligned with IPCC warming ranges

## Step 1: Compare ensmeble distirbution to IPCC range

We begin with SSP2-4.5 because it is the IPCC's recommended scenario for assessing model consistency. It provides a mid-range forcing case suitable for evaluating whether the ensemble aligns with the IPCC's assessed warming uncertainty range.

```{r}
# Extract SSP2-4.5 warming results
ssp_data <- warming_results$`SSP2-4.5`

# Define IPCC Target Values - This information can be found in TABLE 7.SM.4 of IPCC AR6 WGI Ch. 7
ipcc_median <- 1.81  # Median target for SSP2-4.5
ipcc_lower <- 1.24   # 5th percentile target
ipcc_upper <- 2.59   # 95th percentile target

# Compute ensemble percentiles
ensemble_median <- quantile(ssp_data$metric_result, 0.5)
ensemble_lower <- quantile(ssp_data$metric_result, 0.05)
ensemble_upper <- quantile(ssp_data$metric_result, 0.95)

# Display ensemble distribution relative to IPCC Targets
cat("Original Median:", ensemble_median, "\n")
cat("Original 5th Percentile:", ensemble_lower, "\n")
cat("Original 95th Percentile:", ensemble_upper, "\n")
```

After running this, compare the ensemble's median and spread against IPCC targets. This will inform whether filtering is needed and whether the ensemble covers the assessed range sufficiently. 

## Step 2: Filter the ensmeble based on IPCC range

Filter the ensemble to retain runs within (or near) the IPCC WGI-assessed range. 
A small tolerance is applied to prevent over-trimming the ensemble and enough members remain for sub-sampling. 
The allowance values can be adjusted to modify the width of the accepted range. 

```{r}
# Filter ensemble members within the IPCC range
filtered_df <- ssp_data %>%
  filter(metric_result >= (ipcc_lower - 0.3), # adjust allowance as needed to fit IPCC
         metric_result <= (ipcc_upper + 0.08)) # adjust allowance as needed to fit IPCC

# Report the number of retained runs
cat("Number of runs after initial filtering:", nrow(filtered_df), "\n")

```

This step reduces the ensemble to only include runs that are approximately compatible with the IPCC's range. This gives us an initial candidate set of ensemble members for further refinement.

## Step 3: Subsample to optimize IPCC alignment

Subsample from the filtered set to identify an ensemble subset that best matches the IPCC WGI-assessed median and percentiles.

```{r}
# Optimize subset to align with IPCC target values
set.seed(44)  # Ensure reproducibility

best_subset <- NULL # Placeholder to store best subset
best_score <- Inf  # Initialize best score with a large value 

# Loop to search for the best-matching ensemble subset
for (i in 1:1000) {  # Try 1000 different subsets
  
  # Draw a random sample of 6000 ensemble members from the full ensemble
  candidate_subset <- slice_sample(filtered_df, n = 6000, replace = FALSE)

  # Compute percentiles of the candidate subset
  med <- median(candidate_subset$metric_result)
  low <- quantile(candidate_subset$metric_result, 0.05)
  up <- quantile(candidate_subset$metric_result, 0.95)

  # Calculate a simple score based on the absolute deviation from IPCC targets
  score <- abs(med - ipcc_median) + abs(low - ipcc_lower) + abs(up - ipcc_upper)

  # Keep the candidate subset if it outperforms the current best subset
  if (score < best_score) {
    best_score <- score
    best_subset <- candidate_subset
  }
}

# Store the best-matching sibset as the final ensemble
resampled_subset <- best_subset

# Report the final statistics of the selected ensemble
filtered_median <- median(resampled_subset$metric_result)
filtered_lower <- quantile(resampled_subset$metric_result, 0.05)
filtered_upper <- quantile(resampled_subset$metric_result, 0.95)

cat("Filtered Median:", filtered_median, "\n")
cat("Filtered 5th Percentile:", filtered_lower, "\n")
cat("Filtered 95th Percentile:", filtered_upper, "\n")

```

This step identifies the best-matching subset of ensemble members, minimizing deviation from the IPCC WGI-assessed median, 5th, and 95th percentiles. This subset will serve as the IPCC-aligned ensemble for downstream analysis.

```{r}
# Extract the dataset for SSP2-4.5
ssp_data <- warming_results$`SSP2-4.5`

# Define IPCC target values
ipcc_median <- 1.81  
ipcc_lower <- 1.24    
ipcc_upper <- 2.59

# Compute original ensemble statistics
ensemble_median <- median(ssp_data$metric_result)
ensemble_lower <- quantile(ssp_data$metric_result, 0.05)
ensemble_upper <- quantile(ssp_data$metric_result, 0.95)

# Compute absolute deviation from IPCC targets
ssp_data <- ssp_data %>%
  mutate(
    median_diff = abs(metric_result - ipcc_median),
    lower_diff = abs(metric_result - ipcc_lower),
    upper_diff = abs(metric_result - ipcc_upper)
  )

# Assign weights: Lower values get higher weight (better match)
ssp_data <- ssp_data %>%
  mutate(
    weight = 1 / (1 +  3 * abs(metric_result - ipcc_median) + 
                       4 * abs(metric_result - ipcc_lower) + 
                       2 * abs(metric_result - ipcc_upper))
  )

# Normalize weights to sum to 1
ssp_data$weight <- ssp_data$weight / sum(ssp_data$weight)

# Resample based on weights
best_subset <- NULL
best_score <- Inf  # for best score, lower is better

for (i in 1:1000) {  # Try 1000 different subsets
  candidate_subset <- slice_sample(ssp_data, n = 4900, weight_by = weight, replace = F)
  
  # Compute quantiles
  med <- median(candidate_subset$metric_result)
  low <- quantile(candidate_subset$metric_result, 0.05)
  up <- quantile(candidate_subset$metric_result, 0.95)

  # Compute error score based on IPCC targets
  score <- abs(med - ipcc_median) + abs(low - ipcc_lower) + abs(up - ipcc_upper)

  # Keep best match
  if (score < best_score) {
    best_score <- score
    best_subset <- candidate_subset
  }
}

resampled_subset <- best_subset

# Compute final statistics after resampling
filtered_median <- median(resampled_subset$metric_result)
filtered_lower <- quantile(resampled_subset$metric_result, 0.05)
filtered_upper <- quantile(resampled_subset$metric_result, 0.95)

# Print results
cat("Original Median:", ensemble_median, "\n")
cat("Filtered Median:", filtered_median, "\n")
cat("Original 5th Percentile:", ensemble_lower, "\n")
cat("Filtered 5th Percentile:", filtered_lower, "\n")
cat("Original 95th Percentile:", ensemble_upper, "\n")
cat("Filtered 95th Percentile:", filtered_upper, "\n")

```

*Okay this is still not working as intended and its getting a little ridiculous* 
*I think the next thing to try is to have two sets of results:*
*1) A set of results from the low ecs full ensemble that gets us as close as possible to the IPCC ranges without being overly controlling.*
*2) a set if results from the full 15k ensmeble that gets us as close as possible to the IPCC range even if it require forcefully filtering out the ensemble members that have too high a differences from IPCC on the low end*


Now that we have some filtered ensemble that generally aligns with IPCC, extract the run_number from the selected subset.

```{r}
# Extract run numbers from the selected subset
ipcc_aligned_run_numbers <- resampled_subset$run_number

# Print number of selected runs
cat("Number of selected run numbers:", length(ipcc_aligned_run_numbers), "\n")

# check to make sure the run_numbers are identical, if everything done correctly they should be
identical(ipcc_aligned_run_numbers, resampled_subset$run_number)

```
Now that we have these run numbers selected, we want to filter all SSPs using these run_numbers.

Ensure all SSPs use only these identified IPCC-aligned runs: 
```{r}

# Filter each SSP using only the IPCC-aligned run numbers
filtered_results <- lapply(warming_results, function(df) {
  
  df %>% filter(run_number %in% ipcc_aligned_run_numbers)

  })

# Check if all SSPs now have the same number of runs
filtered_run_counts <- sapply(filtered_results, nrow)

print(filtered_run_counts)  # Should be identical across all SSPs

```

### Can skip because all scenarios had the same number of runs to begin with.
Okay unfortunately, even though we used the IPCC-aligned runs, all SSPs still do not have the exact same run_numbers. So we need to identify what the strict common run_numbers are:

<!-- ```{r} -->
<!-- # Find run numbers that exist in ALL SSPs and store as a vector -->
<!-- strict_common_run_vector <- Reduce(intersect,  -->
<!--                                    lapply(filtered_results, function(df) df$run_number)) -->

<!-- strict_common_run_numbers <- data.frame(run_number = strict_common_run_vector) -->

<!-- # Print how many runs are in the strict common set -->
<!-- cat("Number of runs in strict common set:", length(strict_common_run_vector), "\n") -->

<!-- ``` -->

Now that the strict common runs are identified -- 4453 total runs across all SSP scenarios, we will re-filter all SSPs to match exactly:

<!-- ```{r} -->
<!-- # Apply strict filtering so all SSPs have the exact same runs -->
<!-- filtered_results <- lapply(filtered_results, function(df) { -->

<!--   df %>% filter(run_number %in% strict_common_run_vector) -->

<!-- }) -->

<!-- # Verify final run counts -->
<!-- final_run_counts <- sapply(filtered_results, nrow) -->
<!-- print(final_run_counts)  # All numbers should now be identical -->

<!-- ``` -->


## Filtering other SSPs

This is a test to see if the ensemble that produces a good fit for SSP2-4.5 will produce a good fit for the other SSPs. 

I am using the run_numbers for the filtered ensemble from SSP245 to filter other SSPs and then checking their quantiles against IPCC values.

Check to see how the other SSPs align with IPCC when we use this subset of ensembles.

```{r}
lapply(filtered_results, function(df) {
  
  quantile(df$metric_result, c(0.05, 0.50, 0.95))
  
})
```

Now that we have a semi-final set of runs that mostly reproduce WGI assessed warming values, we can filter for parameter sets that result in those ensmeble members.

```{r}
## NOTE that the params have to match the params used to run the model (this can be overlooked so make sure you check)
init_params <- read.csv("data/parameters/initial_parameters_low_ecs_10k.csv", stringsAsFactors = F)

init_params$run_number <- 1:nrow(init_params)

# Filter the parameter set to match the strict common run numbers
IPCC_params <- init_params %>% 
  filter(run_number %in% ipcc_aligned_run_numbers) 

# Convert the run_number column to numeric
IPCC_params$run_number <- as.numeric(IPCC_params$run_number)

# check that the new param run_numbers are the same as `ipcc_aligned_run_numbers` and `resmapled_subset`
nrow(IPCC_params)

# check that the IPCC_params run_numbers are equivalent to the resampled_subset run_numbers
identical(sort(IPCC_params$run_number), sort(resampled_subset$run_number))
```

Check:
```{r}
# Verify that the number of parameter sets matches the filtered ensemble
print(nrow(IPCC_params))  # Should be the same as filtered_results count

print(sapply(filtered_results, nrow))
```

Save these "IPCC" parameters so that we can use them to run a test to see is importing them to run hector will produce the IPCC range.

```{r}
IPCC_params <- IPCC_params %>% 
  select(-run_number)
write.csv(IPCC_params, "output/parameters/IPCC_params.csv", row.names = F)
```

## Visualizations

data prep:
```{r}
gsat_normalized_df <- do.call(rbind, gsat_data_normalized)

gsat_normalized_df <- subset(gsat_normalized_df, year > 1994)

filtered_gsat_data <- lapply(gsat_data_normalized, function(df) {
  
  df %>% filter(run_number %in% ipcc_aligned_run_numbers)
  
})

filtered_gsat_df <- do.call(rbind, filtered_gsat_data)

filtered_gsat_df <- subset(filtered_gsat_df, year > 1994)
```

```{r}

# full_ensemble <- ggplot() +
#   
#   geom_line(data = gsat_normalized_df, 
#             aes(x = year, y = value, group = run_number), 
#             color = "dodgerblue", alpha = 0.5) +
# 
#   geom_line(data = filtered_gsat_df, 
#             aes(x = year, y = value, group = run_number), 
#             color = "salmon", alpha = 0.3) +
#   
#     theme_light() +
#     facet_wrap(~scenario, scales = "free_y")
# 
# full_ensemble

```


### Computing values for full AR6 assessed warming range table:
```{r}
# initialize metric
long_warming_metric <- new_metric(var = 'global_tas', years = 2081:2100, op = median)
mid_warming_metric <- new_metric(var = 'global_tas', years = 2041:2060, op = median)
short_warming_metric <- new_metric(var = 'global_tas', years = 2021:2040, op = median)

metric_list <- list('2081-2100' = long_warming_metric,
                    '2041-2060' = mid_warming_metric,
                    '2021-2040' = short_warming_metric)

# Compute warming metrics using filtered ensemble
warming_result_post <- lapply(names(filtered_gsat_data), function(scenario_name) {
   
   # Copy gsat data (filtered)
   df <- filtered_gsat_data[[scenario_name]]
   
   # Compute metrics for each term length for the current scenario
   metrics_by_term <- lapply(names(metric_list), function(term_length) {

     # Copy metric data
     metric <- metric_list[[term_length]]

     # Compute metrics
     result <- metric_calc(df, metric)

     # Add term_length column
     result$term_length <- term_length

     return(result)
   })

   # rbind term_lengths for the current scenario
   metrics_by_scenario <- do.call(rbind, metrics_by_term)

   # Add column for scenario name
   metrics_by_scenario$scenario <- scenario_name

   return(metrics_by_scenario)
})

# Bind metric results for all scenarios into one data frame
TEST_metrics_df <- do.call(rbind, warming_result_post)
```

Compute warming summary
```{r}
warming_summary <- TEST_metrics_df %>% 
  group_by(scenario, term_length) %>% 
  summarise(
    lower = round(quantile(metric_result, 0.05), 2), 
    central = round(median(metric_result), 2), 
    upper = round(quantile(metric_result, 0.95), 2),
  .groups = "drop")

write.csv(warming_summary, "output/summary_results/warming_summary.csv")

```


Not super informative other than to show that the filtering step largely trims ensemble members from the extremes. But, this is done strategically on my part to trim the distribution asymmetrically to try and get the best alignment with IPCC.

*This figure primarily illustrates how the filtering step trims ensemble members, particularly at the extremes of the distribution. The filtering is applied asymmetrically to optimize alignment with IPCC projections, ensuring that the retained subset better matches the expected warming ranges while preserving internal variability.*

## NEW TODO:

1. Sample ECS with a new prior distribution: mean = 2.7 SD = 0.7 distribution = lognormal
2. run the model with the new ECS samples (and other parameters)
3. Score against history (can't leave this out because of the independent parameter sampling)
4. then try to align results to IPCC range.