---
title: "Warming quantification using IPCC params"
author: "Joe Brown"
date: "2025-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data

Load full result:
```{r}
# load result 
model_result_rerun <- readRDS("output/model_results/update_posterior_result_low_ecs.RDS")

model_result_rerun <- lapply(model_result_rerun, function(df){
  
  no_NAs <- na.omit(df)
  
  return(no_NAs)
  })
```

## Set-up 

Source files in `set-up` folder.

```{r}
source("set-up/source_all.R")

```

## Computing summary stats

```{r, message=FALSE, warning=FALSE}
## subset to only include the variable of interest
gsat_data_rerun <- lapply(model_result_rerun, function(df) {  ## Still by-passing historical scoring step.
  
  subset_data = subset(df, variable == "global_tas")
  
  return(subset_data)
})

## normalize gsat data 
gsat_data_rerun_normalized <- lapply(gsat_data_rerun, function(df) {
  
  normalize_to_reference(df, reference_start = 1995, reference_end = 2014)
  
}) 

## compute summary statistics
summary_rerun <- lapply(names(gsat_data_rerun_normalized), function(scenario_name) {
  
  df <- gsat_data_rerun_normalized[[scenario_name]]
  
  summary_wts <- df %>% 
  group_by(year) %>%
  summarise(
    median = quantile(value, probs = 0.50),
    ci_05 = quantile(value, probs = 0.05),
    ci_10 = quantile(value, probs = 0.10),
    ci_33 = quantile(value, probs = 0.33),
    ci_67 = quantile(value, probs = 0.67),
    ci_90 = quantile(value, probs = 0.90),
    ci_95 = quantile(value, probs = 0.95),
    .groups = "drop"
  ) %>% 
  subset(year > 1849 & year < 2101)
  
  summary_wts$scenario <- scenario_name
  
  return(summary_wts)
  
})

names(summary_rerun) <- names(model_result_rerun)

```

## Computing warming results

### Define metrics of interest

We define a warming metric to give us median warming in the year 2100 give the filtered Hector ensemble.

```{r define metric for end of century warming}
warming_metric <- new_metric(var = "global_tas", years = 2081:2100, op = median)
  
```

### Compute warming results 

Using the `metric_calc` function we compute the warming results.
```{r compute warming metrics}
warming_results_rerun <- lapply(gsat_data_rerun_normalized, function(df) {
  
  metric_result <- metric_calc(df, warming_metric)
  
  return(metric_result)
})
```

*The above code is still not giving me the correct quantile results that should match the values in the Rmd `04_searching_for_IPCC_aligned_ensemble`.

proof:
```{r}
quantile(warming_results_rerun$`SSP2-4.5`$metric_result, probs = c(0.05, 0.5, 0.95))
```

These should equal the filtered quantiles from:

Original Median: 1.730989 
Original 5th Percentile: 1.098986 
Original 95th Percentile: 2.603079 
Filtered Median: 1.749183 
Filtered 5th Percentile: 1.204623 
Filtered 95th Percentile: 2.528252 

But instead it is matching the original quantiles. Therefore something is going wrong between when I calculate the filtered runs and when I compute the "posterior" quantiles.


```{r}
# # initialize metric 
# long_warming_metric <- new_metric(var = 'global_tas', years = 2081:2100, op = median)
# mid_warming_metric <- new_metric(var = 'global_tas', years = 2041:2060, op = median)
# short_warming_metric <- new_metric(var = 'global_tas', years = 2021:2040, op = median)
# 
# metric_list <- list('2081-2100' = long_warming_metric, 
#                     '2041-2060' = mid_warming_metric, 
#                     '2021-2040' = short_warming_metric)
# 
# # compute mid-term warming metrics
# warming_result <- lapply(names(gsat_data_rerun_normalized), function(scenario_name) {
#   
#   # copy gsat data
#   df <- gsat_data_normalized[[scenario_name]]
#   
#   # compute metrics for each term length for current scenario element 
#   metrics_by_term <- lapply(names(metric_list), function(term_length) {
# 
#     # copy metric data
#     metric <- metric_list[[term_length]]
#         
#     # compute metrics
#     result <- metric_calc(df, metric)
#     
#     # add term_length column
#     result$term_length <- term_length
#   
#     return(result)  
#   })  
#   
#   # rbind term_lengths for currrent scenario element
#   metrics_by_scenario <- do.call(rbind, metrics_by_term)
#   
#   # Add column for scenario name
#   metrics_by_scenario$scenario <- scenario_name
#   
#   return(metrics_by_scenario)
# 
# })
# 
# # bind metric result for all scenarios to one df
# metrics_df <- do.call(rbind, warming_result)

```

```{r}
## Wamring median and range

wamring_ranges <- metrics_df %>% 
  group_by(term_length, scenario) %>% 
  summarise(
    lower = quantile(metric_result, probs = c(0.05)), 
    central = quantile(metric_result, probs = c(0.5)), 
    upper = quantile(metric_result, probs = c(0.95)),
  .groups = "drop")

```


