---
title: "Running the model"
author: "Joe Brown"
date: "2025-03-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set-up

Source files in `set-up` folder.

```{r}
source("set-up/source_all.R")
```

# Generate Initial Parameters

Generate initial parameters to be used for model runs.

Load .ini files:
```{r}
# ini directory 
ini_dir <- paste0(system.file("input", package = "hector"), "/")

# read ini files into a list 
ini_list <- list("SSP1-1.9" = paste0(ini_dir, "hector_ssp119.ini"), 
                 "SSP1-2.6" = paste0(ini_dir, "hector_ssp126.ini"), 
                 "SSP2-4.5" = paste0(ini_dir, "hector_ssp245.ini"), 
                 "SSP3-7.0" = paste0(ini_dir, "hector_ssp370.ini"), 
                 "SSP5-8.5" = paste0(ini_dir, "hector_ssp585.ini"))

```

Build perturbed parameter ensemble:
```{r}
# set seed for replication
set.seed(123)

# set sample size
n = 5000

# initiate a core using one of the ini files
param_core <- newcore(ini_list[[3]])

# generate parameters using default sampling
params <- generate_params(core = param_core, draws = n)

# shutdown the core after generating parameters
shutdown(param_core)
```

Sample new ECS distirbution for "low-ECS":
```{r}
# sample new low-ECS distribtion
low_ecs_sample <- rlnorm(n, lognorm(2.7, 0.7)[1], lognorm(2.7, 0.7)[2])

# Add new ECS samples 
params$ECS <- low_ecs_sample

```

# Save Job

Save initial parameter set:
```{r}
params_dir <- here("data", "parameters")

file_name <- "initial_parameters_low_ecs.csv"

write.csv(params, file.path(params_dir, file_name))

```

# Run the model with new "low-ECS" parameters

## Split parameters into separate jobs

Slit parameters into jobs -- will be a list:

```{r}
# splitting params

param_chunks <- split(params, 1:1000)

```

## Set up: Parallel Computing

```{r}

source("set-up/parallel_set_up.R")

```

## Run model using parallel

```{r}
# Run model in parallel mode
result <- parLapply(cl, names(ini_list), function(scenario_name) {
  
  # store data in scenario object
  scenario <- ini_list[[scenario_name]]
  
  # initialize core with the names of the current scenario
  core <- newcore(scenario, name = scenario_name)
  
  # run model for each param chunk for the current core 
  result_list <- lapply(param_chunks, function(chunk) {
    
    iterate_model(core = core, 
                  params = chunk, 
                  save_years = 1800:2200, 
                  save_vars = c("global_tas", 
                                "gmst", 
                                "CO2_concentration", 
                                "ocean_uptake"))
  })
  
  # ensure correct run_number from one chunk to the next
  for (i in 2:length(result_list)) {
    
    max_run_number <- max(result_list[[i - 1]]$run_number)
    
    result_list[[i]]$run_number <- result_list[[i]]$run_number + max_run_number
  }
  
  # bind chunked results into a df for each scenario
  result_return <- do.call(rbind, result_list)
  
  return(result_return)
  
})

# Add names to the list 
names(result) <- names (ini_list)

stopCluster(cl)
```

## Save job

```{r}
result_dir <- here("output", "model_results")

file_name <- "initial_result_low_ecs.RDS"

saveRDS(result, file.path(result_dir, file_name))

```

